---
title: "Base R Programming"
author: "Tyler Hunt"
format: html
editor: visual
---

## Medical Experiment on Blood Pressure

We are creating two vectors. The vector named "pre" holds pre-treatment blood pressure values and "post" holds post-treatment blood pressure values

## Task 1

```{r}
pre <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
post <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

Lets assign common names to both vectors so we know what patient we're looking at.

```{r}
subject_ids <- paste("Subject",1:20,sep="_")
names(pre) <- subject_ids
names(post) <- subject_ids
```

Now that we have names assigned to the patients, lets capture the difference in post vs pre-treatment measures and the average result form these differences. We can simply do this with basic vector math (pre vector - post vector) and calculate the average from this new vector

```{r}
diff <- pre - post
average_diff <- mean(diff)
average_diff
```

Okay, so on average it seems the patients experienced a decrease of 5.65 in blood pressure. Now which patients (and their indices) experience a decrease? Of those that experience a decrease what is the average decrease?

The code below provides the patient identifier (Subject_No.) and the index. In the new vector ("subvector"), we are passing the indicies to only pull patients with a positive decrease.

```{r}
which(diff > 0)
positive_diff_indices <- diff > 0
subvector <- diff[positive_diff_indices]
mean(subvector)
```

We can see from the above that of those patients that have a decreased in blood pressure, the average decrease is 10.6 units

## Task 2

Lets create a data frame consisting of columns for patients, pre-treatments blood pressure, post-treatment blood pressure and the different between pre and post.

```{r}
df <- data.frame(patient = subject_ids,pre_bp = pre ,post_bp = post ,diff_bp =diff)
df
```

From this dataframe, lets only keep rows where diff_bp is greater than 0. We accomplish this by taking a subset of the dataframe and passing a condition to only return records where the difference in blood pressure is greater than 0 (positive decrease)

```{r}
decrease_df <- subset(df,df$diff_bp > 0)
```

Lets add a column that looks at post_bp. If that value is < 120 we want the new column to show TRUE. Otherwise, FALSE.

```{r}
decrease_df$post_less_120 = decrease_df$post_bp < 120
```

Lets print our final result. As a recap, we made the following adjustments to the initial dataframe we created.

1.  Created a subset dataframe that contains only subjects where their "diff_bp" > 0
2.  Add a new column "post_less_120" which is TRUE if "post_bp" \< 120

```{r}
knitr::kable(decrease_df)
```

## Task 3

Lets consider a new set of patients and their data. We will also assign names and replicate some calculations from previous tasks

```{r}
pre_t3 <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_t3 <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
diff_t3 <- pre_t3 - post_t3
subject_ids_t3 <- paste("Subject",1:10,sep="_")

t3_init_df <- data.frame(patient = subject_ids_t3,pre_bp = pre_t3 ,post_bp = post_t3 ,diff_bp = diff_t3, post_less_120 = post_t3 < 120)

t3_df <- subset(t3_init_df, t3_init_df$diff_bp > 0)
```

Lets create a list that includes the dataframes from task 2 (treatment) & task 3 (placebo)

```{r}
list_of_df <- list(treatment = decrease_df, placebo = t3_df)
```

Lets list some different ways we can access the first item within that list. We can use single-square brackets, double-square brackets or reference the name of the first item directly ("treatment" in this case)

```{r}
list_of_df[1]
list_of_df[[1]]
list_of_df["treatment"]
```

Lets see if we can access just the pre_bp column of the placebo dataset. We accomplish this by taking advantage by chaining our $ operator. We are using it to access the "placebo" dataframe and "pre_bp" within that dataframe.

```{r}
list_of_df$placebo$pre_bp
```
