---
title: "Base R Programming"
author: "Tyler Hunt"
format: html
editor: visual
---

## Medical Experiment on Blood Pressure

We are creating two vectors. The vector named "pre" holds pre-treatment blood pressure values and "post" holds post-treatment blood pressure values

## Running Code

```{r}
pre <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
post <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

Lets assign common names to both vectors so we know what patient we're looking at.

```{r}
subject_ids <- paste("Subject",1:20,sep="_")
names(pre) <- subject_ids
names(post) <- subject_ids
```

Now that we have names assigned to the patients, lets capture the difference in post vs pre-treatment measures and the average result form these differences.

```{r}
diff <- pre - post
average_diff <- mean(diff)
```

Okay, so on average it seems the patients experienced a decrease of 5.65 in blood pressure. Now which patients (and their indices) experience a decrease?

```{r}
which(diff > 0)
positive_diff_indices <- diff > 0
subvector <- diff[positive_diff_indices]
```

